---
title: "DS_Kaggle_BikeShare_Prediction_Models"
author: "Divya Sriram"
date: "6/10/2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Decision trees are particulary nice to use when predicting continous outcome variables.

```{r}
# cat("\014")  
setwd("~/Desktop/MIDS/DivyaGitHub/TpT-BikeShareKaggle")

#libraries

library(rpart) #for tree
library(Metrics) #for rmsle
library(party)
library(randomForest)


train_data = read.csv("../TpT-BikeShareKaggle/FeatureEng_Data_Files/train_data.csv", sep = ',')
dev_data = read.csv("../TpT-BikeShareKaggle/FeatureEng_Data_Files/dev_data.csv", sep = ',')
test_data = read.csv("../TpT-BikeShareKaggle/FeatureEng_Data_Files/test_data.csv", sep = ',')

summary(train_data)
```

# (1) RPART MODEL

Using rpart (recursive partitioning and regression trees)

### (1a) RPART Train Data
Let's try use the rpart model to train with our train_data set.

```{r}

# choosing the variables to include in the model
formula_rpart = count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday

# fitting forumula to the model
fit_rpart = rpart(formula_rpart, data=train_data)

# tells us the importance of each variable in the model
fit_rpart

plot(fit_rpart)
text(fit_rpart, use.n=TRUE)

```

According to this model, the most important factor is hour (biggest split).

### (1b) RPART Predict With Dev Data Set

Let's try use the rpart model to predict with our dev_data set. And then we can calculate rmsle to evaluate our model.

```{r}
#dev_data
predict_rpart_dev = predict(fit_rpart, dev_data)

# putting our predictions + hours into dataframe
submit_rpart_dev = data.frame(datetime = dev_data$datetime, count=predict_rpart_dev)

#checking root mean squared log error (like the evaluation in kaggle)
rmsle(dev_data$count, abs(predict_rpart_dev))
```

### (1c) RPART Predict With Test Data Set

Let's try use the rpart model to predict with our test_data set. We'll save the predictions for the test_data set along with the datetime column as a dataframe and convert and save that into a csv file to upload to kaggle.

```{r}
#test_data
predict_rpart_test = predict(fit_rpart, test_data)

# putting our predictions + hours into dataframe
submit_rpart_test = data.frame(datetime = test_data$datetime, count=predict_rpart_test)

# writing the dataframe to a csv file --> submit to kaggle
write.csv(submit_rpart_test, file="../TpT-BikeShareKaggle/Submission_Files/rpart/submit_rpart_test_v3.csv",row.names=FALSE)

```

# (2) PARTY MODEL

### (2a) PARTY Train Data

Let's try use the party model to train with our train_data set.

Using party (recursive partitioning and regression trees)

```{r}

# choosing the variables to include in the model
formula_ctree = count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday

#fitting forumula to the model
fit_ctree = ctree(formula_ctree, data=train_data)

#tells us the importance of each variable in the model
fit_ctree

plot(fit_ctree)
```


According to this model, the most important factor is temp (biggest split).

### (2b) PARTY Predict With Dev Data Set

Let's try use the party model to predict with our dev_data set. And then we can calculate rmsle to evaluate our model.

```{r}
#dev_data
predict_ctree_dev = predict(fit_ctree, dev_data)

# putting our predictions + hours into dataframe
submit_ctree_dev = data.frame(datetime = dev_data$datetime, count=predict_ctree_dev)

#checking root mean squared log error (like the evaluation in kaggle)
rmsle(dev_data$count, abs(predict_ctree_dev))
```


### (2c) PARTY Predict With Test Data Set

Let's try use the party model to predict with our test_data set. We'll save the predictions for the test_data set along with the datetime column as a dataframe and convert and save that into a csv file to upload to kaggle.

```{r}
#test_data
predict_ctree_test = predict(fit_ctree, test_data)

# putting our predictions + hours into dataframe
submit_ctree_test = data.frame(datetime = test_data$datetime, count=predict_ctree_test)

# writing the dataframe to a csv file --> submit to kaggle
write.csv(submit_ctree_test, file="../TpT-BikeShareKaggle/Submission_Files/party/submit_ctree_test_changedseasonsfewvariables.csv",row.names=FALSE)

```

# (3) RANDOM FORESTS MODEL

### Random Forests Train Data
```{r}
# choosing the variables to include in the model
formula_rf = count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday

# fitting forumula to the model
rf_model = randomForest(formula_rf, data=train_data, ntree = 250)

# tells us the importance of each variable in the model
print(rf_model)
```

### Random Forests Predict With Dev Data Set

```{r}
#dev_data
predict_rf_dev = predict(rf_model, dev_data)

# putting our predictions + hours into dataframe
submit_rf_dev = data.frame(datetime = dev_data$datetime, count=predict_rf_dev)

#checking root mean squared log error (like the evaluation in kaggle)
rmsle(dev_data$count, abs(predict_rf_dev))

```

### Random Forests Predict With Test Data Set

```{r}
#test_data
predict_rf_test = predict(rf_model, test_data)

# putting our predictions + hours into dataframe
submit_rf_test = data.frame(datetime = test_data$datetime, count=predict_rf_test)

# writing the dataframe to a csv file --> submit to kaggle
write.csv(submit_rf_test, file="../TpT-BikeShareKaggle/Submission_Files/randomforest/submit_rf_test_250trees3var_UNchangedseasons.csv",row.names=FALSE)
```

-----------------------------------------------------------------
# SUBMISSIONS RECORDS:

1. submit_rpart_test_v1.csv: 0.90215
- variables all (count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday)
- unchanged seasons

2. submit_ctree_test_changedseasonsfewvariables.csv: 0.67175
- not all variables
- changed seasons

3. submit_ctree_test_changedseasons.csv: 0.63706
- variables all (count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday)
- changed seasons

4. submit_rf_test_250trees3var.csv: 0.60693
- variables all (count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday)
- 250 trees
- 3 variables
- changed seasons

5. submit_rf_test_250trees3var_removedwindspeed.csv : 0.60693
- variables (no windspeed) (count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday)
- 250 trees
- 3 variables
- changes seasons

6. submit_rf_test_250trees3var_UNchangedseasons.csv : (0.59960)
- variables all (count ~ hour + temp + humidity + season + weather + dayofweek + windspeed + month + workingday)
- 250 trees
- 3 variables
- UNchanged seasons






